---
title: "Getting_data_week2"
output: html_document
date: "2022-09-14"
---
##SQL
Connection and listing databases
dbConnect(MySQL(), user = , db = database name,host = )#open the connection
dbGetQuery()
dbDisconnect()
dbListTables()
dbListField()


dbGetQuery(database name,  "select count(*) from name of the table")
dbReadTable(database name , "name of the table")
dbSendQuery(name of database, "select * from name of table where misMatches between 1 and 3")
fetch()
quantile()
dbClearResult()
dim()

##HDF5
library(fhdf5)
h5createFile()
h5createGroup()
h5ls() h5write() h5read()

##Reading data from web
url("")
readLines()
close()
htmlCode()

##parsing with XML
library(XML)
url <- "..."
html <- htmlTreeParse(url, useInternalNodes = T)
xpathSApply(html, "//title". xmlValue)
xpathSAplly(html, "//td[@id='col-citedby']", xmlValue)

library(httr)
html2 =GET(url)
content = content(html2, as = "text")
parsedHtml = htmlParse(content2, asText = TRUE)
xpathSAPlly(parsedHtml,"//title", xmlValue)
Accessing websites with passwords
pg1 = Get("", authenticate("user","passwd"))

google = handle("")
pg1 = GET(handle = google, path = "/")
pg2 = GET(handle = google, path = "search")


From APIs (Application programming interfaces)

```{r}
library(httr)

# 1. Find OAuth settings for github:
#    http://developer.github.com/v3/oauth/
oauth_endpoints("github")

# 2. To make your own application, register at
#    https://github.com/settings/developers. Use any URL for the homepage URL
#    (http://github.com is fine) and  http://localhost:1410 as the callback url
#
#    Replace your key and secret below.
myapp <- oauth_app("github",
  key = "56b637a5baffac62cad9",
  secret = "8e107541ae1791259e9987d544ca568633da2ebf"
)

# 3. Get OAuth credentials
github_token <- oauth2.0_token(oauth_endpoints("github"), myapp)

# 4. Use API
gtoken <- config(token = github_token)
req <- GET("https://api.github.com/rate_limit", gtoken)
stop_for_status(req)
content(req)

# OR:
req <- with_config(gtoken, GET("https://api.github.com/rate_limit"))
stop_for_status(req)
content(req)
```

Question 2
```{r}
library("sqldf")
setwd('/Users/dengxiaoyan/Documents/Rstudio/Getting and Cleaning Data/')
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
f <- file.path(getwd(), "ss06pid.csv")
download.file(url, f)
acs <- data.table::data.table(read.csv(f))

# Answer: 
query1 <- sqldf("select pwgtp1 from acs where AGEP < 50")

```

Question 4
```{r}
library(httr)
connection <- url( "http://biostat.jhsph.edu/~jleek/contact.html")
htmlCode <- readLines(connection)
close(connection)
c(nchar(htmlCode[10]), nchar(htmlCode[20]), nchar(htmlCode[30]), nchar(htmlCode[100]))

```
Question 5
```{r}
x <- read.fwf(
  file=url("https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for"),
  skip=4,
  widths=c(12, 7, 4, 9, 4, 9, 4, 9, 4))
sum(x[,4])
```